### 2022-01-07 RunMSE/HCR_MSY/MSEPlots.Rmd

---
title: "Application of Huynh 2020 Interim Analysis methods to other stocks: investigating if stocks tend toward MSY with HCR_MSY harvest control rule."
author: "Nikolai Klibansky"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    extra_dependencies:
      longtable: null
      xcolor: null
urlcolor: blue
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())
knitr::opts_chunk$set(comment=NA,echo=FALSE,message=FALSE,warning=FALSE,fig.height=7,fig.width=6)

library(Hmisc)
library(knitr)
#library(latex2exp) # lets you use TeX() function to convert latex equations to R expressions for plotting
library(SAMtool)
library(stringr)
library(bamExtras)

source('fn/iMP_NK.R')
# Compute U (harvest rate) from F
F2U <- function(F){1 - exp(-F)}

# Median without NA
median_noNA <- function(x){median(x,na.rm=TRUE)}
```

```{r user}

OMNames_user <- c("MSE_BlackSeaBass","MSE_VermilionSnapper","MSE_RedPorgy")#,
#                    #"OM_Tilefish","OM_Cobia","OM_SnowyGrouper",
#                    #"OM_capelin", "OM_POP", "OM_vs",
#   "OM_Albacore","OM_AlmacoJack", "OM_BandedMorwong"
#                    )
# OMNames_user <- local({
#   a <- gsub(".rds","",list.files("MSE_obj"))
#   b <- gsub("MSE","OM",a)
#   b
# })

OMNames_user <- gsub("MSE","OM",OMNames_user)
scenNames_user <-  c("base","nr"#,#"hs","hd",
                     #"dep","lf",
                     #"epiM"
                     )
OMscenNames_user <- sort(apply(expand.grid(OMNames_user,scenNames_user),1,function(x){paste(x,collapse="_")}))
scenNames_key <- c("base"="Base","nr"="no rec error","hs"="Hyper-stable","hd"="Hyper-deplete",
                   "dep"="Depleted","lf"="Lightly fished","epiM"="Episodic M")
MPNamesSub_user <- c("SCA_1","SCA_10","iMP_avg_10","MP_PerfectMSY")#,"iMP_avg_10","iMP_buffer_10")
MPNamesSub_user_legtext <- c("Annual assessment","Fixed TAC (10)","Averaged Index (10)","Perfect MSY")#,"Buffered Index (10)")
```

```{r plot_LH function, eval=TRUE}
plot_LH <- function(OM,...){
maxage <- OM@maxage
agebins <- 0:maxage
lenbins <- vb_len(Linf=OM@Linf[1],K=OM@K[1],t0=OM@t0[1],a=agebins)
Plenbins <- lenbins/max(lenbins)
Mat_age <- OM@cpars$Mat_age[1,,1] # Maturity at age
V <- OM@cpars$V[1,,1] # Vulnerability (selectivity) at age

plot(agebins,Plenbins,ylim=c(0,1),type="l",lwd=3,...)
points(agebins,Mat_age,type="l",lwd=3,lty=2)
points(agebins,V,type="l",lwd=3,lty=3)
}
```

```{r plot_MSE_ts_Huynh function}

plot_MSE_ts_Huynh <- function(
myvarName,
myvarDenomExpr = NULL,
ylabel = myvarName,
FUN = mean_noNA,
scenNames,
OMNames,
MPNamesSub = c("SCA_10","SCA_1","iMP_avg_10","iMP_buffer_10"),
MPNamesSub_legtext=MPNamesSub,
legtext_o = c(1,2,3,4), # order for plotting legend text
MSEoutNames = paste0(rep(gsub("OM","MSE",OMNames),each=length(scenNames)),
                     "_",
                     rep(scenNames,length(OMNames))
                     ),
colLabs = NULL,
rowLabs = scenNames,

mar=c(1,1,1,1),
cols = setNames(c("black","red","deepskyblue2","green3"),MPNamesSub),
ltys = setNames(c(1,2,1,2),MPNamesSub),
lwds = setNames(c(3,3,3,3),MPNamesSub),
types = setNames(c("l","l","l","l"),MPNamesSub),
xlabel = "Management year",
xlim = NULL,
ylim = NULL,
assessmentYears = seq(10,40,by=10),
x.at = NULL,
y.at = NULL,
# leg.x=NULL,
# leg.y =NULL,
legend_panel=NULL,
legend_cex= 1,
legend_x="bottomleft",
legend_inset=c(0,-0.3),
label_axes="all", # "all" labels all panels, "leftbottom" labels y axes on the left and x axes on the bottom panels
hline=1 # values to pass into abline(h=hline)
){
if(is.null(colLabs)){
  colLabs <- local({
  a <- gsub("OM_","",OMNames) # Replace OM_
  b <- str_replace_all(a,"_[A-Za-z]*$","") # Replace underscore and letters following
  c <- str_replace_all(b,"(?<=[a-z])(?=[A-Z])"," ")
  unique(c)
  })
}

xlim_user <- xlim
ylim_user <- ylim
x.at_user <- x.at
y.at_user <- y.at

colLabsAll = as.character(matrix(c(colLabs,rep("",length(MSEoutNames)-length(colLabs))),ncol=length(colLabs),byrow=TRUE))
rowLabsAll = as.character(matrix(c(rep("",length(MSEoutNames)-length(rowLabs)),rowLabs),nrow=length(rowLabs),byrow=FALSE))

nrow <- length(rowLabs)
ncol <- length(colLabs)
npanel <- nrow*ncol

par(mfcol=c(nrow,ncol),mgp=c(1,0.2,0),mar=mar,oma=c(6,4,3,3),tck=-0.01,xpd=FALSE,xaxs="i",yaxs="i")

for(outName_i in MSEoutNames){
topPanels <- (length(scenNames)*(seq_along(OMNames)-1)+1)
bottomPanels <- length(scenNames)*seq_along(OMNames)
leftPanels <- seq_along(scenNames)
rightPanels <- tail(seq_along(MSEoutNames),length(scenNames))

i <- which(MSEoutNames==outName_i)
out_i <- get(outName_i)
MPNames <- out_i@MPs
if(!is.expression(myvarName)){
myvar <- slot(out_i,myvarName)  
}else{
myvar <- eval(myvarName) 
}

if(is.null(legend_panel)){
legend_panel <- bottomPanels[1]
}

if(!is.null(myvarDenomExpr)){
myvarDenom <- eval(myvarDenomExpr)
myvar <- myvar/myvarDenom # Scale myvar
}

if(!is.null(dim(myvar))){
myvarsub <- myvar[,match(MPNamesSub,MPNames),,drop=FALSE]
myvarsubStat <- apply(myvarsub,2,function(x){apply(x,2,FUN)})
}else{
myvarsubStat <- matrix(rep(myvar,length(MPNamesSub)),ncol=length(MPNamesSub))  
}
dimnames(myvarsubStat) <- list(1:nrow(myvarsubStat),MPNamesSub)

xaxt_i <- ifelse(i%in%bottomPanels,"n","n")
yaxt_i <- ifelse(i%in%leftPanels,"n","n")

if(is.null(ylim_user)){
  ylim <- range(myvarsubStat,na.rm=TRUE)
}
if(is.null(xlim_user)){
  xlim <- range(as.numeric(rownames(myvarsubStat)),na.rm=TRUE)
}

matplot(as.numeric(rownames(myvarsubStat)),myvarsubStat,
        col=cols,lty=ltys,lwd=lwds,type=types,
        xaxt=xaxt_i,yaxt=yaxt_i,
        xlim=xlim,ylim=ylim,xlab="",ylab="")
#title(main=colLabsAll[i],line=1)

grid(nx=NA,ny=NULL)
abline(v=assessmentYears,lty="1111",lwd=2)
if(!is.null(hline)){
abline(h=hline)
}

if(is.null(x.at_user)){
  x.at <- pretty(par("usr")[1:2])
}
if(is.null(y.at_user)){
  y.at <- pretty(par("usr")[3:4])
}

if(i%in%topPanels){
col.ct <- which(topPanels==i)
mtext(colLabsAll[i],side=3,line=1,outer=FALSE,font=2)
}

if(i%in%rightPanels){
row.ct <- which(rightPanels==i)
rowcenter <-par("usr")[3]+diff(par("usr")[3:4])/2

text(x=par("usr")[2]*1.05,y=rowcenter,rowLabsAll[i],
     srt=-90,xpd=NA,font=2)
}

if(label_axes=="all"){
  x.at.i <- x.at
  axis(side=1,at=x.at.i)
  y.at.i <- y.at
  axis(side=2,at=y.at.i)
}
if(label_axes=="leftbottom"){
  if(i%in%bottomPanels){
    x.at.i <- x.at
    if(i>bottomPanels[1]){x.at.i <- x.at[-1]}
    axis(side=1,at=x.at.i)
  }
  if(i%in%leftPanels){
    y.at.i <- y.at
    if(i>leftPanels[1]){y.at.i <- y.at[1:(length(y.at)-1)]}
    axis(side=2,at=y.at.i)
  }
}

if(i==legend_panel){
   legend(legend_x,
          inset=legend_inset,
       horiz=TRUE,bty="n",
       legend=MPNamesSub_legtext[legtext_o],
       col=cols[legtext_o],lty=ltys[legtext_o],lwd=lwds[legtext_o],
       cex=legend_cex,
       xpd=NA
       )

}
}

mtext(ylabel,side=2,line=1.5,outer=TRUE)
mtext(xlabel,side=1,line= par("mgp")[1],outer=TRUE)

}
```

```{r plot_results function}
plot_results <- function(OMNames,plots=c("SSB/SSBmsy","F/Fmsy","Catch/MSY","TAC/MSY")){
  speciesNames <- gsub("OM_","",OMNames)

  # SSB/SSBmsy
  if("SSB/SSBmsy"%in%plots){    
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = "SB_SBMSY",
      ylabel = expression(mean~~SSB/SSBmsy),
      FUN = mean,
      ylim=c(0,2),
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # F/Fmsy      
  if("F/Fmsy"%in%plots){
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = "F_FMSY",
      ylabel = expression(mean~~F/Fmsy),
      FUN = mean,
      ylim=c(0,4),
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # Catch/MSY
  if("Catch/MSY"%in%plots){    
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = "Catch",
      myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears),drop=FALSE]),
      ylabel = expression(mean~~Catch/MSY),
      FUN = mean,
      ylim=c(0,2),
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # TAC/MSY
  if("TAC/MSY"%in%plots){ 
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = "TAC",
      myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears),drop=FALSE]),
      ylabel = "TAC/MSY",
      FUN = mean,
      ylim=c(0,4),
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # Catch/TAC
  if("Catch/TAC"%in%plots){ 
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = "Catch",
      myvarDenomExpr = expression(slot(out_i,"TAC")),
      ylabel = expression(mean~~Catch/TAC),
      FUN = mean,
      ylim=c(0,1.1),
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # SSB
  if("SSB"%in%plots){ 
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = "SSB",
      ylabel = expression(mean~~SSB),
      FUN = mean,
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # VB
  if("VB"%in%plots){     
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = "VB",
      ylabel = expression(mean~~VB),
      FUN = mean,
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # F
  if("F"%in%plots){     
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = "FM",
      ylabel = expression(mean~~F),
      FUN = mean,
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # Catch
  if("Catch"%in%plots){ 
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user, 
      myvarName = "Catch",
      ylabel = expression(mean~~Catch),
      FUN = mean,
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # TAC
  if("TAC"%in%plots){     
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = "TAC",
      FUN = mean,
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user] 
    )
  }

  # MSY
  if("MSY"%in%plots){     
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears),drop=FALSE]),
      FUN = mean,
      ylabel="MSY",
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # SSBMSY
  if("SSBMSY"%in%plots){ 
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = expression(slot(out_i,"RefPoint")$SSBMSY[,,-(1:out_i@nyears),drop=FALSE]),
      FUN = mean,
      ylabel="SSBMSY",
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # FMSY
  if("FMSY"%in%plots){     
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = expression(slot(out_i,"RefPoint")$FMSY[,,-(1:out_i@nyears),drop=FALSE]),
      FUN = mean,
      ylabel="FMSY",
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }

  # UMSY
  if("UMSY"%in%plots){     
    plot_MSE_ts_Huynh(
      OMNames= OMNames, scenNames=scenNames_user,
      myvarName = expression(F2U(slot(out_i,"RefPoint")$FMSY[,,-(1:out_i@nyears),drop=FALSE])),
      FUN = mean,
      ylabel="UMSY",
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user]
    )
  }  
}
```

```{r data, results='hide'}
# Read in data for original operating models
for(file.i in list.files("OM")){
  assign(gsub(".rds","",file.i),readRDS(file.path("OM",file.i)))
}

# Read in data for modified operating models
for(file.i in list.files("OM_modified")){
  assign(gsub(".rds","",file.i),readRDS(file.path("OM_modified",file.i)))
}

if(is.null(OMNames_user)){
OMNames <- gsub(".rds","",list.files("OM"))
}else{
OMNames <- OMNames_user
}

OM_nsim <- c()
for(OM_Name_i in OMNames){
  i <- which(OMNames==OM_Name_i)
  OM_i <- get(OM_Name_i)
  nsim_i <- OM_i@nsim
  OM_nsim <- c(OM_nsim,nsim_i)
}

nsim_text <- combine_words(unique(OM_nsim),and=" or ")

# Read in MSE results
MSE_res_names <- gsub(".rds","",list.files("MSE_obj"))
#MSE_res_names <- gsub(".rds","",list.files("MSE_obj")[grepl("OM_",list.files("MSE_obj"))])
for(file.i in list.files("MSE_obj")){
  obj_name <- gsub(".rds","",file.i)
  print(obj_name)
  assign(obj_name,readRDS(file.path("MSE_obj",file.i)))
}

```

```{r plot data1, fig.cap=paste0("\\label{fig:lifeHist} Life history (growth and maturity) and vulnerability schedules at age used in the operating models for ",combine_words(speciesNames),". Growth is expressed as mean length-at-age relative to that at the maximum age. Compare with Huynh et al (2020) Figure 1."),eval=FALSE}
# Figure 1 from Huynh et al (2020)

# NOTE: These won't always work if the life history values aren't in the right slots.

par(mfrow=c(length(OMNames),1),mgp=c(1,0.2,0),mar=c(1.5,2,1.5,1),oma=c(1,2,1,1),tck=-0.01)
for(OM_Name_i in OMNames){
i <- which(OMNames==OM_Name_i)  
OM_i <- get(OM_Name_i)
speciesName_i <- gsub("OM_","",OM_Name_i)
plot_LH(OM_i,xlab="",ylab="",main=speciesName_i)
if(i==1){
  legend("bottomright",legend=c("Rel. length","Maturity","Vulnerability"),lwd=c(3,3,3),lty=c(1,2,3),bty="n")
}
if(i==length(OMNames)){
  mtext("Value",2,0,outer = TRUE)
}
}
```

```{r plot results,eval=TRUE}
# ct <- 1
# while(ct<=length(OMscenNames_user)){
#   OMscenNames_i <- as.character(na.omit(OMscenNames_user[ct+0:2]))
#   plot_results(OMNames=OMscenNames_i)
#   ct <- ct+3
# }

  plot_results(OMNames=OMNames_user)

```