### 2022-01-07 RunMSE/SEFSC/MSEPlots.Rmd

---
title: "Application of Huynh 2020 Interim Analysis methods to assessed stocks from the US South Atlantic"
author: "Nikolai Klibansky"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    extra_dependencies:
      longtable: null
      xcolor: null
urlcolor: blue
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())
knitr::opts_chunk$set(comment=NA,echo=FALSE,message=FALSE,warning=FALSE,fig.height=7,fig.width=6)

library(Hmisc)
library(knitr)
library(latex2exp) # lets you use TeX() function to convert latex equations to R expressions for plotting
library(SAMtool)
library(stringr)
library(bamExtras)

source('fn/iMP_NK.R')
# Compute U (harvest rate) from F
F2U <- function(F){1 - exp(-F)}

# Median without NA
median_noNA <- function(x){median(x,na.rm=TRUE)}
```

```{r user}
OMNames_user <- c("OM_BlackSeaBass","OM_VermilionSnapper","OM_RedPorgy")#,
                   #"OM_Tilefish","OM_Cobia","OM_SnowyGrouper",
                   #"OM_capelin", "OM_POP", "OM_vs"
                   #)
scenNames_user <-  c("base",#"hs","hd",
                     "dep","lf",
                     "epiM"
                     )
scenNames_key <- c("base"="Base","hs"="Hyper-stable","hd"="Hyper-deplete",
                   "dep"="Depleted","lf"="Lightly fished","epiM"="Episodic M")
MPNamesSub_user <- c("SCA_10","SCA_1","iMP_avg_10","iMP_buffer_10")
MPNamesSub_user_legtext <- c("Fixed TAC (10)","Annual assessment","Averaged Index (10)","Buffered Index (10)")

# OMNames_user <- c("OM_BlackSeaBass","OM_SnowyGrouper","OM_VermilionSnapper")
# OMNames_user <- c("OM_RedPorgy","OM_Tilefish","OM_Cobia")
```

```{r plot_LH function}
plot_LH <- function(OM,...){
maxage <- OM@maxage
agebins <- 0:maxage
lenbins <- vb_len(Linf=OM@Linf[1],K=OM@K[1],t0=OM@t0[1],a=agebins)
Plenbins <- lenbins/max(lenbins)
Mat_age <- OM@cpars$Mat_age[1,,1] # Maturity at age
V <- OM@cpars$V[1,,1] # Vulnerability (selectivity) at age

plot(agebins,Plenbins,ylim=c(0,1),type="l",lwd=3,...)
points(agebins,Mat_age,type="l",lwd=3,lty=2)
points(agebins,V,type="l",lwd=3,lty=3)
}
```

```{r plot_MSE_ts_Huynh function}

plot_MSE_ts_Huynh <- function(
myvarName,
myvarDenomExpr = NULL,
ylabel = myvarName,
FUN = mean_noNA,
scenNames,
OMNames,
MPNamesSub = c("SCA_10","SCA_1","iMP_avg_10","iMP_buffer_10"),
MPNamesSub_legtext=MPNamesSub,
legtext_o = c(2,3,4,1), # order for plotting legend text
MSEoutNames = paste0(rep(gsub("OM","MSE",OMNames),each=length(scenNames)),
                     "_",
                     rep(scenNames,length(OMNames))
                     ),
colLabs = str_replace_all(gsub("OM_","",OMNames),"(?<=[a-z])(?=[A-Z])"," "),
rowLabs = scenNames,

mar=c(1,1,1,1),
cols = setNames(c("red","black","deepskyblue2","deepskyblue2"),MPNamesSub),
ltys = setNames(c(1,1,1,2),MPNamesSub),
lwds = setNames(c(3,3,3,3),MPNamesSub),
types = setNames(c("l","l","l","l"),MPNamesSub),
xlabel = "Management year",
xlim = NULL,
ylim = NULL,
assessmentYears = seq(10,40,by=10),
x.at = NULL,
y.at = NULL,
# leg.x=NULL,
# leg.y =NULL,
legend_panel=NULL,
legend_cex= 1,
legend_x="bottomleft",
legend_inset=c(0,-0.6),
label_axes="all", # "all" labels all panels, "leftbottom" labels y axes on the left and x axes on the bottom panels
hline=1 # values to pass into abline(h=hline)
){
xlim_user <- xlim
ylim_user <- ylim
x.at_user <- x.at
y.at_user <- y.at

colLabsAll = as.character(matrix(c(colLabs,rep("",length(MSEoutNames)-length(colLabs))),ncol=length(colLabs),byrow=TRUE))
rowLabsAll = as.character(matrix(c(rep("",length(MSEoutNames)-length(rowLabs)),rowLabs),nrow=length(rowLabs),byrow=FALSE))

nrow <- length(rowLabs)
ncol <- length(colLabs)
npanel <- nrow*ncol

par(mfcol=c(nrow,ncol),mgp=c(1,0.2,0),mar=mar,oma=c(6,4,3,3),tck=-0.01,xpd=FALSE,xaxs="i",yaxs="i")

for(outName_i in MSEoutNames){
topPanels <- (length(scenNames)*(seq_along(OMNames)-1)+1)
bottomPanels <- length(scenNames)*seq_along(OMNames)
leftPanels <- seq_along(scenNames)
rightPanels <- tail(seq_along(MSEoutNames),length(scenNames))

i <- which(MSEoutNames==outName_i)
out_i <- get(outName_i)
MPNames <- out_i@MPs
if(!is.expression(myvarName)){
myvar <- slot(out_i,myvarName)  
}else{
myvar <- eval(myvarName) 
}

legend_panel <- bottomPanels[1]

if(!is.null(myvarDenomExpr)){
myvarDenom <- eval(myvarDenomExpr)
myvar <- myvar/myvarDenom # Scale myvar
}

if(!is.null(dim(myvar))){
myvarsub <- myvar[,match(MPNamesSub,MPNames),]
myvarsubStat <- apply(myvarsub,2,function(x){apply(x,2,FUN)})
}else{
myvarsubStat <- matrix(rep(myvar,length(MPNamesSub)),ncol=length(MPNamesSub))  
}
dimnames(myvarsubStat) <- list(1:nrow(myvarsubStat),MPNamesSub)

xaxt_i <- ifelse(i%in%bottomPanels,"n","n")
yaxt_i <- ifelse(i%in%leftPanels,"n","n")

if(is.null(ylim_user)){
  ylim <- range(myvarsubStat,na.rm=TRUE)
}
if(is.null(xlim_user)){
  xlim <- range(as.numeric(rownames(myvarsubStat)),na.rm=TRUE)
}

matplot(as.numeric(rownames(myvarsubStat)),myvarsubStat,
        col=cols,lty=ltys,lwd=lwds,type=types,
        xaxt=xaxt_i,yaxt=yaxt_i,
        xlim=xlim,ylim=ylim,xlab="",ylab="")
#title(main=colLabsAll[i],line=1)

grid(nx=NA,ny=NULL)
abline(v=assessmentYears,lty="1111",lwd=2)
if(!is.null(hline)){
abline(h=hline)
}

if(is.null(x.at_user)){
  x.at <- pretty(par("usr")[1:2])
}
if(is.null(y.at_user)){
  y.at <- pretty(par("usr")[3:4])
}

if(i%in%topPanels){
col.ct <- which(topPanels==i)
mtext(colLabsAll[i],side=3,line=1,outer=FALSE,font=2)
}

if(i%in%rightPanels){
row.ct <- which(rightPanels==i)
rowcenter <-par("usr")[3]+diff(par("usr")[3:4])/2

text(x=par("usr")[2]*1.05,y=rowcenter,rowLabsAll[i],
     srt=-90,xpd=NA,font=2)
}

if(label_axes=="all"){
  x.at.i <- x.at
  axis(side=1,at=x.at.i)
  y.at.i <- y.at
  axis(side=2,at=y.at.i)
}
if(label_axes=="leftbottom"){
  if(i%in%bottomPanels){
    x.at.i <- x.at
    if(i>bottomPanels[1]){x.at.i <- x.at[-1]}
    axis(side=1,at=x.at.i)
  }
  if(i%in%leftPanels){
    y.at.i <- y.at
    if(i>leftPanels[1]){y.at.i <- y.at[1:(length(y.at)-1)]}
    axis(side=2,at=y.at.i)
  }
}

if(i==legend_panel){
   legend(legend_x,
          inset=legend_inset,
       horiz=TRUE,bty="n",
       legend=MPNamesSub_legtext[legtext_o],
       col=cols[legtext_o],lty=ltys[legtext_o],lwd=lwds[legtext_o],
       cex=legend_cex,
       xpd=NA
       )

}
}

mtext(ylabel,side=2,line=1.5,outer=TRUE)
mtext(xlabel,side=1,line= par("mgp")[1],outer=TRUE)

}
```

```{r data}
# Read in data for original operating models
for(file.i in list.files("OM")){
  assign(gsub(".rds","",file.i),readRDS(file.path("OM",file.i)))
}

# Read in data for modified operating models
for(file.i in list.files("OM_modified")){
  assign(gsub(".rds","",file.i),readRDS(file.path("OM_modified",file.i)))
}

if(is.null(OMNames_user)){
OMNames <- gsub(".rds","",list.files("OM"))
}else{
OMNames <- OMNames_user
}

OM_nsim <- c()
for(OM_Name_i in OMNames){
  i <- which(OMNames==OM_Name_i)
  OM_i <- get(OM_Name_i)
  nsim_i <- OM_i@nsim
  OM_nsim <- c(OM_nsim,nsim_i)
}

nsim_text <- combine_words(unique(OM_nsim),and=" or ")

# Read in MSE results
MSE_res_names <- gsub(".rds","",list.files("MSE_obj"))
#MSE_res_names <- gsub(".rds","",list.files("MSE_obj")[grepl("OM_",list.files("MSE_obj"))])
for(file.i in list.files("MSE_obj")){
  assign(gsub(".rds","",file.i),readRDS(file.path("MSE_obj",file.i)))
}

# # Merge sets of MSE results
# for(i in  OMNames_user){
#   for(j in scenNames_user){
#     OMscenName <- paste0(i,"_",j)
#     MSEName <- gsub("OM","MSE",OMscenName)
#     a <- mget(MSE_res_names[grepl(OMscenName,MSE_res_names)])
#     b <- do.call(merge_MSE, args=a)
#     assign(MSEName,b)
#   }
# }
# 
# rm(list=MSE_res_names)
# MSE_res_names <- gsub("OM","MSE",apply(expand.grid(OMNames_user,scenNames_user),1,function(x){paste(x,collapse ="_")}))

```

```{r,eval=TRUE}
OMNames <- c("OM_BlackSeaBass","OM_VermilionSnapper","OM_RedPorgy")
speciesNames <- gsub("OM_","",OMNames)
```

```{r plot data1, fig.cap=paste0("\\label{fig:lifeHist} Life history (growth and maturity) and vulnerability schedules at age used in the operating models for ",combine_words(speciesNames),". Growth is expressed as mean length-at-age relative to that at the maximum age. Compare with Huynh et al (2020) Figure 1."),eval=TRUE}
# Figure 1 from Huynh et al (2020)

par(mfrow=c(length(OMNames),1),mgp=c(1,0.2,0),mar=c(1.5,2,1.5,1),oma=c(1,2,1,1),tck=-0.01)
for(OM_Name_i in OMNames){
i <- which(OMNames==OM_Name_i)  
OM_i <- get(OM_Name_i)
speciesName_i <- gsub("OM_","",OM_Name_i)
plot_LH(OM_i,xlab="",ylab="",main=speciesName_i)
if(i==1){
  legend("bottomright",legend=c("Rel. length","Maturity","Vulnerability"),lwd=c(3,3,3),lty=c(1,2,3),bty="n")
}
if(i==length(OMNames)){
  mtext("Value",2,0,outer = TRUE)
}
}
```

```{r plot results1,eval=TRUE}

# SSB/SSBmsy
plot_MSE_ts_Huynh(
  OMNames= OMNames, scenNames=scenNames_user,
  myvarName = "SB_SBMSY",
  ylabel = expression(mean~~SSB/SSBmsy),
  FUN = mean,
  ylim=c(0,2),
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user]
)

# F/Fmsy
plot_MSE_ts_Huynh(
  OMNames= OMNames, scenNames=scenNames_user,
  myvarName = "F_FMSY",
  ylabel = expression(mean~~F/Fmsy),
  FUN = mean,
  ylim=c(0,2),
  MPNamesSub = MPNamesSub_user,
  rowLabs=scenNames_key[scenNames_user]
)

# C/MSY
plot_MSE_ts_Huynh(
  OMNames= OMNames, scenNames=scenNames_user,
  myvarName = "Catch",
  myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
  ylabel = expression(mean~~Catch/MSY),
  FUN = mean,
  ylim=c(0,2),
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user]
)

# TAC/MSY
plot_MSE_ts_Huynh(
  OMNames= OMNames, scenNames=scenNames_user,
  myvarName = "TAC",
  myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
  ylabel = "TAC/MSY",
  FUN = mean,
  ylim=c(0,4),
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user]
)

# C/TAC
plot_MSE_ts_Huynh(
  OMNames= OMNames, scenNames=scenNames_user,  
  myvarName = "Catch",
  myvarDenomExpr = expression(slot(out_i,"TAC")),
  ylabel = expression(mean~~C/TAC),
  FUN = mean,
  ylim=c(0,1.1),
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user]
)

# SSB
plot_MSE_ts_Huynh(
  OMNames= OMNames, scenNames=scenNames_user,
  myvarName = "SSB",
  ylabel = expression(mean~~SSB),
  FUN = mean,
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user]
)

# VB
plot_MSE_ts_Huynh(
OMNames= OMNames, scenNames=scenNames_user,
myvarName = "VB",
ylabel = expression(mean~~VB),
FUN = mean,
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user]
)

# F
plot_MSE_ts_Huynh(
OMNames= OMNames, scenNames=scenNames_user,  
myvarName = "FM",
ylabel = expression(mean~~F),
FUN = mean,
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user]
)

# Catch
plot_MSE_ts_Huynh(
OMNames= OMNames, scenNames=scenNames_user,  
myvarName = "Catch",
ylabel = expression(mean~~Catch),
FUN = mean,
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user]
)

# TAC
plot_MSE_ts_Huynh(
OMNames= OMNames, scenNames=scenNames_user,  
myvarName = "TAC",
FUN = mean,
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user] 
)

# MSY
plot_MSE_ts_Huynh(
OMNames= OMNames, scenNames=scenNames_user,
myvarName = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
FUN = mean,
ylabel="MSY",
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user]
)

# SSBMSY
plot_MSE_ts_Huynh(
OMNames= OMNames, scenNames=scenNames_user,
myvarName = expression(slot(out_i,"RefPoint")$SSBMSY[,,-(1:out_i@nyears)]),
FUN = mean,
ylabel="SSBMSY",
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user]
)

# FMSY
plot_MSE_ts_Huynh(
OMNames= OMNames, scenNames=scenNames_user,
myvarName = expression(slot(out_i,"RefPoint")$FMSY[,,-(1:out_i@nyears)]),
FUN = mean,
ylabel="FMSY",
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user]
)

# UMSY
plot_MSE_ts_Huynh(
OMNames= OMNames, scenNames=scenNames_user,
myvarName = expression(F2U(slot(out_i,"RefPoint")$FMSY[,,-(1:out_i@nyears)])),
FUN = mean,
ylabel="UMSY",
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user]
)

```