### 2022-03-16 RunMSE/SEFSC/MSEPlots.Rmd

---
title: "Application of Huynh 2020 Interim Analysis methods to assessed stocks from the US South Atlantic"
author: "Nikolai Klibansky"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    extra_dependencies:
      longtable: null
      xcolor: null
      pdflscape: null
urlcolor: blue
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())
knitr::opts_chunk$set(comment=NA,echo=FALSE,message=FALSE,warning=FALSE,fig.width=8,fig.height=10)

library(Hmisc)
library(knitr)
library(latex2exp) # lets you use TeX() function to convert latex equations to R expressions for plotting
library(SAMtool)
library(stringr)
library(bamExtras)

#source('fn/iMP.R')
# Compute U (harvest rate) from F
F2U <- function(F){1 - exp(-F)}

# Median without NA
median_noNA <- function(x){median(x,na.rm=TRUE)}
```

```{r user}
# Should individual PDFs be produced (or just plot in Rmd output)
plot_pdf <- TRUE

OMNames_user <- c("OM_BlackSeaBass",
                  "OM_VermilionSnapper",
                  "OM_RedPorgy",
                   #"OM_Tilefish","OM_Cobia",
                  "OM_SnowyGrouper"#,
                  #"OM_GagGrouper")#,
                   #"OM_capelin", "OM_POP", "OM_vs"
                   )
OMNames_huynh <- c("OM_CapelinHuynh", "OM_PacificOceanPerchHuynh", "OM_VermilionSnapperHuynh")

scenNames_user1 <-  c("base","dep","lf","epiM","ubias","tachi","rc")
scenNames_user2 <-  c("base","nolag","minerr","minrecdev","perfobs","ucvhi","ucvlo","vtiv"
                        )

# All scenarios
scenNames_user <- unique(c(scenNames_user1,scenNames_user2))

# Alternative realities
# These scenarios are considered equally plausible, for the purposes of
# averaging results together. Most other scenarios could be considered sensitivities.
scenNames_user_alt <- c("base","dep","lf","epiM","ubias","tachi","rc")

scenNames_key <- c("base"="Base",
                   "nolag"="No data lag",
                   "hs"="Hyper-stable","hd"="Hyper-deplete",
                   "dep"="Depleted","lf"="Lightly fished","epiM"="Episodic M",
                   "rc"="Regime\nchange", "ubias"="Index\nbiased","ucvhi"="Index\nHigh CV",
                   "ucvlo"="Index\nLow CV",
                   "perfobs"="Perfect observation",
                   "minerr"="Minimal error",
                   "minrecdev"="Minimal rec devs",
                   "tachi"="TAC 1.25MSY",
                   "vtiv"="V time invariant"
                   )

MPNamesAll <-         c(#"AvC",
                        "CC1",
                        #"DCAC","DBSRA","SPMSY",
                        "SCA_1","SCA_5","SCA_10",
                        "iMP_avg_5","iMP_avg_10",
                        "iMP_buffer_5","iMP_buffer_10",
                        "pMP_5","pMP_10"
                        )

MPNamesAll_key <- c("AvC"="average catch",
                    "CC1"="average catch (last 5 yr)",
                    "DCAC"= "depletion corrected average catch",
                    "DBSRA" = "depletion based stock reduction analysis",
                    "SPMSY" = "surplus production model",
                    "SCA_1" = "SCA 1 yr",
                    "SCA_5" = "SCA 5 yr",
                    "SCA_10"= "SCA 10 yr",
                    "iMP_avg_5" = "avg. index (5 yr)",
                    "iMP_avg_10" = "avg. index (10 yr)",
                    "iMP_buffer_5" = "buffer index (5 yr)",
                    "iMP_buffer_10" = "buffer index (10 yr)",
                    "pMP_5" = "projection (5 yr)",
                    "pMP_10" = "projection (10 yr)"
                        )
MPNamesAll_legtext <- MPNamesAll

MPNamesSub_user <- c("SCA_10","SCA_5","SCA_1","iMP_avg_10"#,"iMP_buffer_10"
                     )
MPNamesSub_user_legtext <- c("Fixed TAC (10)","Fixed TAC (5)","Annual assessment","Averaged Index (10)"#,"Buffered Index (10)"
                             )

# OMNames_user <- c("OM_BlackSeaBass","OM_SnowyGrouper","OM_VermilionSnapper")
# OMNames_user <- c("OM_RedPorgy","OM_Tilefish","OM_Cobia")
```

## Built in performance metrics (PMs)

* `P10`	  Probability B > 0.1 BMSY
* `P50`	  Probability B > 0.5 BMSY
* `P100`  Probability B > BMSY
* `PNOF`  Probability F < FMSY
* `LTY`	  Probability Long-Term Yield > 0.5 Relative Yield
* `STY`	  Probability Short-Term Yield > 0.5 Relative Yield
* `AAVY`	Probability AAVY < 0.2 (Average Annual Variability in Yield)
* `AAVE`	Probability AAVE < 0.2 (Average Annual Variability in Effort)
* `Yield`	Average Yield (relative to Reference Yield)

- can easily vary reference values and years
- PMs compute probabilities across years for each simulation and return mean values across simulations (but we could recompute them however we want from the raw MSE objects)

```{r functions}
# color function like rainbow() but without yellow (or indigo)
ROGBP <- colorRampPalette(c("red","orange","green","blue","purple"))
# Makes an ROGBP function with an alpha channel
ROGBP_alpha <- function(alpha){
  x <- colorRampPalette(c(
    rgb(1,0,0,alpha), # red
    rgb(1,0.5,0.25,alpha), # orange
    rgb(0,1,0,alpha), # green 
    rgb(0,0,1,alpha), # blue
    rgb(0.5,0,0.5,alpha)), # purple
    alpha=TRUE)
  return(x)
}

# Turn hexadecimal color opaque
makeOpaque <-      function(hexcolor){gsub(".{2}$","FF",hexcolor)}

P75 <- function(...,Ref=0.75){P50(...,Ref=Ref)}

AAVY10 <- function(MSEobj = NULL, Ref = 0.1, Yrs = NULL){AAVY(MSEobj = MSEobj, Ref = Ref, Yrs = Yrs)}

```

```{r plot_LH function}
plot_LH <- function(OM,...){
maxage <- OM@maxage
agebins <- 0:maxage
lenbins <- vb_len(Linf=OM@Linf[1],K=OM@K[1],t0=OM@t0[1],a=agebins)
Plenbins <- lenbins/max(lenbins)
styr_proj <- OM@nyears+1
Mat_age <- apply(OM@cpars$Mat_age[,,styr_proj],2,mean) # Maturity at age
V <- apply(OM@cpars$V[,,styr_proj],2,mean) # Vulnerability (selectivity) at age

plot(agebins,Plenbins,ylim=c(0,1),type="l",lwd=3,...)
points(agebins,Mat_age,type="l",lwd=3,lty=2)
points(agebins,V,type="l",lwd=3,lty=3)
}
```

```{r plot_MSE_ts function}
# Time series plots

plot_MSE_ts <- function(
myvarName="SSB",
myvarDenomExpr = NULL,
ylabel = myvarName,
FUN = mean,
scenNames,
OMNames,
MPNamesSub = NULL,
MPNamesSub_legtext=MPNamesSub,
MPNamesRef = NULL, # Name of reference MP
FUNRef = function(myvarsubStat,MPNamesRef){ # Function for comparing MPs with reference MP
  (myvarsubStat-myvarsubStat[,MPNamesRef])/myvarsubStat[,MPNamesRef]},
legtext_o = NULL, # order for plotting legend text
MSEoutNames = paste0(rep(gsub("OM","MSE",OMNames),each=length(scenNames)),
                     "_",
                     rep(scenNames,length(OMNames))
                     ),
colLabs = str_replace_all(gsub("OM_","",OMNames),"(?<=[a-z])(?=[A-Z])"," "),
rowLabs = scenNames,
sims = NULL, # which simulation runs should be used to generate the plot when plot_type="sims"?
plot_type="huynh", # "huynh" plots time series by MP like in the Huynh et al paper
                   # "boot" plots times series of bootstrapped data  with bamExtras::plot_boot_vec
                   # "phase" plots phase plots of F/Fmsy versus SSB/SSBmsy
                   # "tradeoff" plots tradeoff plots by MP
                   # "addind" plot additional indices from AddInd
                   # "sims" plot time series for individual sim runs
proyear_phase=NULL,
PMx = Yield, # Performance metric function x (x-axis) for tradeoff plots)
PMy = P75, # Performance metric function y (y-axis) for tradeoff plots)
mar=c(1,1,1,1),
oma=c(6,4,3,3),
mgp=c(1.5,0.2,0),
cols = setNames(ROGBP_alpha(0.5)(length(MPNamesSub)),MPNamesSub),
ltys = setNames(rep(1,length(MPNamesSub)),MPNamesSub),
pchs = setNames(rep(16,length(MPNamesSub)),MPNamesSub),
lwds = setNames(rep(3,length(MPNamesSub)),MPNamesSub),
types = setNames(rep("l",length(MPNamesSub)),MPNamesSub),
xlabel = "Management year",
main_line = 1, # Line for main column labels
xlim = NULL,
ylim = NULL,
assessmentYears=NULL,
x.at = NULL,
y.at = NULL,
# leg.x=NULL,
# leg.y =NULL,
legend_panel=NULL,
legend_cex= 1,
legend_x="bottomleft",
legend_inset=c(0,-0.5),
label_axes="all", # "all" labels all panels, "leftbottom" labels y axes on the left and x axes on the bottom panels
hline=1 # values to pass into abline(h=hline)
){
if(is.null(legtext_o)){
  legtext_o <- seq_along(MPNamesSub)
}

xlim_user <- xlim
ylim_user <- ylim
x.at_user <- x.at
y.at_user <- y.at

colLabsAll = as.character(matrix(c(colLabs,rep("",length(MSEoutNames)-length(colLabs))),ncol=length(colLabs),byrow=TRUE))
rowLabsAll = as.character(matrix(c(rep("",length(MSEoutNames)-length(rowLabs)),rowLabs),nrow=length(rowLabs),byrow=FALSE))

dim1 <- length(rowLabs)
dim2 <- length(colLabs)

if(plot_type=="addind"){
  mfcol <- c(dim2,dim1)}else{
  mfcol <- c(dim1,dim2)
}

par(mfcol=mfcol,mgp=mgp,mar=mar,oma=oma,tck=-0.01,xpd=FALSE,xaxs="r",yaxs="r")

### LOOP THROUGH MSE OBJECTS
for(outName_i in MSEoutNames){
topPanels <- (length(scenNames)*(seq_along(OMNames)-1)+1)
bottomPanels <- length(scenNames)*seq_along(OMNames)
leftPanels <- seq_along(scenNames)
rightPanels <- tail(seq_along(MSEoutNames),length(scenNames))

i <- which(MSEoutNames==outName_i)
out_i <- get(outName_i)
MPNames <- out_i@MPs

if(is.null(sims)){
  sims <- 1:out_i@nsim
}
if(!is.expression(myvarName)){
  myvar <- slot(out_i,myvarName)  
}else{
  myvar <- eval(myvarName) 
}

nyears_i <- out_i@nyears
proyears_i <- out_i@proyears
LHyear_i <- out_i@Hist@Data@LHYear # Last historic year
histYears <- sort(LHyear_i - ((1:nyears_i)-1))
projYears_i <- LHyear_i+1:proyears_i

legend_panel <- bottomPanels[1]

if(!is.null(myvarDenomExpr)){
myvarDenom <- eval(myvarDenomExpr)
myvar <- myvar/myvarDenom # Scale myvar
}

xaxt_i <- ifelse(i%in%bottomPanels,"n","n")
yaxt_i <- ifelse(i%in%leftPanels,"n","n")

MPNamesSubix <- match(MPNamesSub,MPNames)

if(plot_type=="huynh"){
if(!is.null(dim(myvar))){
myvarsub <- myvar[,match(MPNamesSub,MPNames),]
myvarsubStat <- apply(myvarsub,2,function(x){apply(x,2,FUN)})
}else{
myvarsubStat <- matrix(rep(myvar,length(MPNamesSub)),ncol=length(MPNamesSub))  
}
dimnames(myvarsubStat) <- list(1:nrow(myvarsubStat),MPNamesSub)

if(!is.null(MPNamesRef)){
myvarsubStat <- FUNRef(myvarsubStat,MPNamesRef)
}

if(is.null(ylim_user)){
  ylim <- range(myvarsubStat,na.rm=TRUE)
}
if(is.null(xlim_user)){
  xlim <- range(projYears_i,na.rm=TRUE)
}

matplot(projYears_i#as.numeric(rownames(myvarsubStat))
        ,myvarsubStat,
        col=cols,lty=ltys,lwd=lwds,type=types,
        xaxt=xaxt_i,yaxt=yaxt_i,
        xlim=xlim,ylim=ylim,xlab="",ylab="")
#title(main=colLabsAll[i],line=1)

grid(nx=NA,ny=NULL)
abline(v=assessmentYears,lty="1111",lwd=2)
}
    if(plot_type=="sims"){
      cols <-  ROGBP(length(sims))
      ltys <-  1:length(MPNamesSub)
      pchs <-  NA
      lwds <-  3
      types <- "l"

      myvarsub <- myvar[sims,match(MPNamesSub,MPNames),,drop=FALSE]
      dimnames(myvarsub) <- list(sims,MPNamesSub,1:proyears_i)

      if(!is.null(MPNamesRef)){
        myvarsub <- FUNRef(myvarsub,MPNamesRef)
      }

      if(is.null(ylim_user)){
        ylim <- range(myvarsub,na.rm=TRUE)
      }
      if(is.null(xlim_user)){
        xlim <- range(projYears_i,na.rm=TRUE)
      }

      for(MP_j in MPNamesSub){
      j <- match(MP_j,MPNamesSub)
      myvarsub_j <- myvarsub[,j,]
      lty <- ltys[j]
      if(j==1){
      matplot(projYears_i,
              t(myvarsub_j), 
              col=cols,lty=lty,lwd=lwds,type=types,
              xaxt=xaxt_i,yaxt=yaxt_i,
              xlim=xlim,ylim=ylim,xlab="",ylab="")
      #title(main=colLabsAll[i],line=1)

      grid(nx=NA,ny=NULL)
      abline(v=assessmentYears,lty=2,lwd=1)
      }else{
        matpoints(t(myvarsub_j), 
                col=cols,lty=lty,lwd=lwds,type=types
                )        
      }
    }
    }

if(plot_type=="addind"){
  PPDYears_i <- local({
    a <- c(histYears,projYears_i)
    a[-length(a)]
  })
if(is.null(assessmentYears)){
assessmentYears <- apply(do.call(cbind,lapply(out_i@PPD[[MPNamesSub]]@Misc,FUN=function(x){x1 <- x$diagnostic; unlist(lapply(x1,FUN=function(xx){xx$Year}))})),1,median)
}

if(!is.null(dim(myvar))){
myvarsub <- myvar
#myvarsubStat <- apply(myvarsub,2,function(x){apply(x,2,FUN)})
}else{
  warning("is.null(dim(myvar) == TRUE. No option currently available for this situation.")
}
dimnames(myvarsub) <- list("sim"=1:dim(myvarsub)[1],"AddInd"=1:dim(myvarsub)[2],"Year"=PPDYears_i)
if(is.null(ylim_user)){
  ylim <- range(myvarsub,na.rm=TRUE)
}
if(is.null(xlim_user)){
  xlim <- range(PPDYears_i,na.rm=TRUE)
}
# n_AddInd <- ncol(myvarsubStat)
# cols <-  ROGBP_alpha(0.5)(n_AddInd)
# matplot(PPDYears_i,myvarsubStat,
#         col=cols,lty=ltys,lwd=lwds,type=types,
#         xaxt=xaxt_i,yaxt=yaxt_i,
#         xlim=xlim,ylim=ylim,xlab="",ylab="")
#title(main=colLabsAll[i],line=1)
n_AddInd <- dim(myvarsub)[2]
col_AddInd <- ROGBP(n_AddInd)
for(j in 1:n_AddInd){

plot_boot_vec(myvarsub[,j,],
              col_lines = col_AddInd[j],
              add=ifelse(j!=1,TRUE,FALSE),
              xlim=xlim,ylim=ylim)

}
grid(nx=NA,ny=NULL)
abline(v=assessmentYears,lty=2,lwd=1)
}

if(plot_type=="boot"){
  if(is.null(ylim_user)){
  ylim <- range(myvar,na.rm=TRUE)
}
  bamExtras::plot_boot_vec(myvar,ylim=ylim)
  abline(v=out_i@nyears,lwd=2,lty=2) # Plot vertical line at end of historical period
}

if(plot_type=="phase"){
ltys <-  setNames(rep(0,length(MPNamesSub)),MPNamesSub)
# pchs <-  setNames(rep(1,length(MPNamesSub)),MPNamesSub)
lwds <-  setNames(rep(0,length(MPNamesSub)),MPNamesSub)
types <- setNames(rep("p",length(MPNamesSub)),MPNamesSub)  
xlabel <- "F/Fmsy"  
ylabel <- "SSB/SSBmsy"

if(is.null(proyear_phase)){
  proyear_phase <- out_i@proyears
}
Bstatus <- out_i@SB_SBMSY[,MPNamesSubix,proyear_phase]
Fstatus <- out_i@F_FMSY[,MPNamesSubix,proyear_phase]

matplot(Fstatus,Bstatus,
        col=cols,lty=ltys,lwd=lwds,pch=pchs,type=types,
        xaxt=xaxt_i,yaxt=yaxt_i,
        xlab="",ylab="",xlim=xlim,ylim=ylim
        )
points(apply(Fstatus,2,median),
       apply(Bstatus,2,median),pch=21,cex=3,lwd=2,col="black",bg=makeOpaque(cols))

  abline(v=1,lty=3)
  abline(h=1,lty=3)

}

if(plot_type=="tradeoff"){
  out_i <- get(outName_i)

PMx_out <- PMx(out_i)
PMy_out <- PMy(out_i)

plot(PMx_out@Mean[MPNamesSubix],PMy_out@Mean[MPNamesSubix],pch=16,col=cols,
     xaxt=xaxt_i,yaxt=yaxt_i,
     xlab="",ylab="",
     xlim=xlim,ylim=ylim
     )

# legend("top",legend=MPs,col=MPcols,pch=16,bty="n")

ltys <-  setNames(rep(0,length(MPNamesSub)),MPNamesSub)
lwds <-  setNames(rep(0,length(MPNamesSub)),MPNamesSub)
types <- setNames(rep("p",length(MPNamesSub)),MPNamesSub)  
xlabel <- PMx_out@Caption 
ylabel <- PMy_out@Caption

}

if(!is.null(hline)){
abline(h=hline)
}

if(is.null(x.at_user)){
  x.at <- pretty(par("usr")[1:2])
}
if(is.null(y.at_user)){
  y.at <- pretty(par("usr")[3:4])
}

if(i%in%topPanels){
col.ct <- which(topPanels==i)
mtext(colLabsAll[i],side=3,line=main_line,outer=FALSE,font=2)
}

if(i%in%rightPanels){
row.ct <- which(rightPanels==i)
rowcenter <-par("usr")[3]+diff(par("usr")[3:4])/2
text(x=par("usr")[2]+diff(par("usr")[1:2])*0.1
     ,y=rowcenter,rowLabsAll[i],
     srt=-90,xpd=NA,font=2)
}

if(label_axes=="all"){
  x.at.i <- x.at
  axis(side=1,at=x.at.i)
  y.at.i <- y.at
  axis(side=2,at=y.at.i)
}
if(label_axes=="leftbottom"){
  if(i%in%bottomPanels){
    x.at.i <- x.at
    if(i>bottomPanels[1]){x.at.i <- x.at[-1]}
    axis(side=1,at=x.at.i,outer=TRUE)
  }
  if(i%in%leftPanels){
    y.at.i <- y.at
    if(i>leftPanels[1]){y.at.i <- y.at[1:(length(y.at)-1)]}
    axis(side=2,at=y.at.i,outer=TRUE)
  }
}

if(i==legend_panel&plot_type%in%c("huynh","phase","tradeoff")){
   legend(legend_x,
          inset=legend_inset,
       horiz=TRUE,bty="n",
       legend=MPNamesSub_legtext[legtext_o],
       col=makeOpaque(cols)[legtext_o],lty=ltys[legtext_o],lwd=lwds[legtext_o],pch=pchs[legtext_o],
       cex=legend_cex,
       xpd=NA,
       x.intersp=0.5
       )

}
}

mtext(ylabel,side=2,line=par("mgp")[1],outer=TRUE)
mtext(xlabel,side=1,line= par("mgp")[1],outer=TRUE)

}
```

```{r plot_MSE_dw function}
# Dot and whisker plots

plot_MSE_dw <- function(
myvarName,
myvarDenomExpr = NULL,
ylabel = myvarName,
FUN = cv,  # MSEtool::cv
scenNames,
OMNames,
MPNamesSub = c("SCA_10","SCA_1","iMP_avg_10","iMP_buffer_10"),
MPNamesSub_legtext=MPNamesSub,
legtext_o = c(3,2,1,4), # order for plotting legend text
MSEoutNames = paste0(rep(gsub("OM","MSE",OMNames),each=length(scenNames)),
                     "_",
                     rep(scenNames,length(OMNames))
                     ),
colLabs = str_replace_all(gsub("OM_","",OMNames),"(?<=[a-z])(?=[A-Z])"," "),
rowLabs = scenNames,

mar=c(0,0,0,0),
oma=c(10,4,3,3),
mgp=c(1,0.3,0),
cols = setNames(ROGBP(length(MPNamesSub)),MPNamesSub),
ltys = setNames(rep(1,length(MPNamesSub)),MPNamesSub),
lwds = setNames(rep(1,length(MPNamesSub)),MPNamesSub),
pchs = setNames(rep(16,length(MPNamesSub)),MPNamesSub),
types = setNames(rep("l",length(MPNamesSub)),MPNamesSub),
textlabpos = rep("lo",length(MSEoutNames)),#c(rep("lo",6),rep("up",7),"lo",rep("up",4)),
xlabel = "Management procedure",
xlabel_line=9,
xlim = NULL,
ylim = c(0,0.6),
x.at = NULL,
y.at = NULL,
# leg.x=NULL,
# leg.y =NULL,
# legend_panel=NULL,
# legend_cex= 1,
# legend_x="bottomleft",
# legend_inset=c(0,-0.6),
label_axes="leftbottom", # "all" labels all panels, "leftbottom" labels y axes on the left and x axes on the bottom panels
hline=1 # values to pass into abline(h=hline)
){

xlim_user <- xlim
ylim_user <- ylim
x.at_user <- x.at
y.at_user <- y.at

colLabsAll = as.character(matrix(c(colLabs,rep("",length(MSEoutNames)-length(colLabs))),ncol=length(colLabs),byrow=TRUE))
rowLabsAll = as.character(matrix(c(rep("",length(MSEoutNames)-length(rowLabs)),rowLabs),nrow=length(rowLabs),byrow=FALSE))

nrow <- length(rowLabs)
ncol <- length(colLabs)

par(mfcol=c(nrow,ncol),mgp=mgp,mar=mar,oma=oma,tck=-0.01,xpd=FALSE,xaxs="i",
    yaxs="r",lend="butt")

for(outName_i in MSEoutNames){
topPanels <- (length(scenNames)*(seq_along(OMNames)-1)+1)
bottomPanels <- length(scenNames)*seq_along(OMNames)
leftPanels <- seq_along(scenNames)
rightPanels <- tail(seq_along(MSEoutNames),length(scenNames))

i <- which(MSEoutNames==outName_i)
out_i <- get(outName_i)
MPNames <- out_i@MPs
if(!is.expression(myvarName)){
myvar <- slot(out_i,myvarName)  
}else{
myvar <- eval(myvarName) 
}

legend_panel <- bottomPanels[1]

if(!is.null(myvarDenomExpr)){
myvarDenom <- eval(myvarDenomExpr)
myvar <- myvar/myvarDenom # Scale myvar
}

if(!is.null(dim(myvar))){
myvarsub <- myvar[,match(MPNamesSub,MPNames),]
myvarsubStat <- apply(myvarsub,2,function(x){apply(x,2,FUN)})
}else{
myvarsubStat <- matrix(rep(myvar,length(MPNamesSub)),ncol=length(MPNamesSub))  
}
dimnames(myvarsubStat) <- list(1:nrow(myvarsubStat),MPNamesSub)

xaxt_i <- ifelse(i%in%bottomPanels,"n","n")
yaxt_i <- ifelse(i%in%leftPanels,"n","n")

if(is.null(xlim_user)){
  xlim <- c(0,length(MPNamesSub)+1)
}

################################

Statmed <- apply(myvarsubStat,2,median)
Statmedtext <- round(Statmed,2)
StatQ1 <- apply(myvarsubStat,2,quantile,probs=0.25,na.rm=TRUE)
StatQ3 <- apply(myvarsubStat,2,quantile,probs=0.75,na.rm=TRUE)

xaxt_i <- ifelse(i%in%bottomPanels,"n","n")
yaxt_i <- ifelse(i%in%leftPanels,"n","n")
x <- seq_along(MPNamesSub)
plot(x,Statmed,
        col=cols,pch=pchs,type="n",bg="white",
        xaxt=xaxt_i,yaxt=yaxt_i,
        xlim=xlim,ylim=ylim,xlab="",ylab="")
arrows(x0=x,y0=StatQ1,y1=StatQ3,
       col=cols,lty=ltys,lwd=lwds,length=0
       )
points(x,Statmed,
        col=cols,pch=pchs,type="p",bg="white",cex=1.2,
        xaxt=xaxt_i,yaxt=yaxt_i,
        xlim=xlim,ylim=ylim,xlab="",ylab="")

if(textlabpos[i]=="lo"){text(x=x,y=rep(par("usr")[3]),labels=Statmedtext,srt=90,adj=c(-0.1,0.5),font=2)}
if(textlabpos[i]=="up"){text(x=x,y=rep(par("usr")[4]),labels=Statmedtext,srt=90,adj=c(1.2,0.5),font=2)}

################################

if(i%in%topPanels){
col.ct <- which(topPanels==i)
mtext(colLabsAll[i],side=3,line=1,outer=FALSE,font=2)
}

if(i%in%rightPanels){
row.ct <- which(rightPanels==i)
rowcenter <-par("usr")[3]+diff(par("usr")[3:4])/2

text(x=par("usr")[2]*1.05,y=rowcenter,rowLabsAll[i],
     srt=-90,xpd=NA,font=2)
}

if(label_axes=="all"){
  x.at.i <- x.at
  axis(side=1,at=x.at.i)
  y.at.i <- y.at
  axis(side=2,at=y.at.i)
}
if(label_axes=="leftbottom"){
  if(i%in%bottomPanels){
    x.at.i <- x.at
    if(i>bottomPanels[1]){x.at.i <- x.at[-1]}
    # axis(side=1,at=x.at.i)
    axis(side=1,at=x,labels=MPNamesSub_legtext,las=2)
  }
  if(i%in%leftPanels){
    y.at.i <- y.at
    if(i>leftPanels[1]){y.at.i <- y.at[1:(length(y.at)-1)]}
    axis(side=2,at=y.at.i)
  }
}

# if(i==legend_panel){
#    legend(legend_x,
#           inset=legend_inset,
#        horiz=TRUE,bty="n",
#        legend=MPNamesSub_legtext[legtext_o],
#        col=cols[legtext_o],lty=ltys[legtext_o],lwd=lwds[legtext_o],
#        cex=legend_cex,
#        xpd=NA
#        ) 
# 
# }
}

mtext(ylabel,side=2,line=1.5,outer=TRUE)
mtext(xlabel,side=1,line=xlabel_line,outer=TRUE)

}
```

```{r plot_MSE_sims function}
plot_MSE_sims <- function(MSE, sims=NULL, MPs=NULL, proyears_ix=NULL,
              varNames=NULL, rec_age=0,
              type="proj", # period to plot: "proj", or "histproj"
              title_main = NULL,
              vline=NULL,vline_lty=3,
              ...){
MSEName <- deparse(substitute(MSE))
if(is.null(varNames)){
  varNames <- list("proj"=    c("Catch","FM","F_FMSY","recruit","SSB","SB_SBMSY"),
                   "histproj"=c("Catch","FM","F_FMSY","recruit","SSB","SB_SBMSY"))[[type]]
}

nyears <- MSE@nyears
proyears <- MSE@proyears
LHyear <- MSE@Hist@Data@LHYear # Last historic year
histYears <- sort(LHyear - ((1:nyears)-1))
projYears <- LHyear+1:proyears

if(is.null(sims)){sims <- 1:3}  
if(is.null(MPs)){MPs <- MSE@MPs[1]}
if(is.null(proyears_ix))    {proyears_ix <- 1:proyears}

MPn <- match(MPs,MSE@MPs)

nr <- ceiling(sqrt(length(varNames)))
nc <- ceiling(length(varNames)/nr)
par(mfcol=c(nr,nc),mar=c(2,3,1,1),oma=c(0,0,2,0),mgp=c(1,0.2,0),tck=-0.01)

for(varName_i in varNames){
  for(MPn_j in  MPn){
    MPn_j_ix <- match(MPn_j,MPn)
    if(type=="proj"){
      years_plot <- projYears
      if(varName_i=="recruit"){
        N <- Reduce("+",apply(MSE@Misc$extended$N,5,FUN=function(x){x},simplify = FALSE))
        var_i <- N[,rec_age+1,,]
        years_ix <-   tail(1:dim(var_i)[3],MSE@proyears)
        # drop MP dimension but not sims
        varSub_i <- var_i[sims,MPn_j,years_ix,drop=FALSE]
        varSub_i <- matrix(varSub_i,nrow=dim(varSub_i)[1],ncol=dim(varSub_i)[3]) 
      }else{
        var_i <- slot(MSE,varName_i)
        varSub_i <- var_i[sims,MPn_j,proyears_ix,drop=FALSE]
        varSub_i <- matrix(varSub_i,nrow=dim(varSub_i)[1],ncol=dim(varSub_i)[3]) 
      }
    }

    if(type=="histproj"){
      years_plot <- c(histYears, projYears)
      if(varName_i=="recruit"){
        N <- Reduce("+",apply(MSE@Misc$extended$N,5,FUN=function(x){x},simplify = FALSE))
        var_i <- N[,rec_age+1,,]
      }
      if(varName_i%in%c("F","FM")){
        FM <- apply(MSE@Misc$extended[["FM"]],c(1,3,4),max) # Max F over ages and areas
        var_i <- FM
      }
      if(varName_i=="F_FMSY"){
        FM <- apply(MSE@Misc$extended[["FM"]],c(1,3,4),max) # Max F over ages and areas
        FMSY <- MSE@RefPoint$FMSY
        var_i <- FM/FMSY
      }
      if(varName_i=="SB_SBMSY"){
        SSB <- apply(MSE@Misc$extended[["SSB"]],c(1,3,4),sum) # Sum SSB over ages and areas
        SSBMSY <- MSE@RefPoint$SSBMSY
        var_i <- SSB/SSBMSY
      }
      if(!varName_i%in%c("F","FM","recruit","F_FMSY","SB_SBMSY")){
        var_i <- Reduce("+",apply(MSE@Misc$extended[[varName_i]],5,FUN=function(x){x},simplify = FALSE))
        var_i <- Reduce("+",apply(var_i,2,FUN=function(x){x},simplify = FALSE))
      }
      #varSub_i <- var_i[sims,MPn_j,] # Drops dimension to be matrix
        varSub_i <- var_i[sims,MPn_j,,drop=FALSE]
        varSub_i <- matrix(varSub_i,nrow=dim(varSub_i)[1],ncol=dim(varSub_i)[3]) 
    }
    if(match(MPn_j,MPn)==1){
      matplot(years_plot,t(varSub_i),type="l",
              col=ROGBP(length(sims)),xlab="",ylab=varName_i, lty=MPn_j_ix,...)
    }else{
      matpoints(years_plot,t(varSub_i),type="l",
                col=ROGBP(length(sims)),lty=MPn_j_ix,...)
    }    
  }
  if(match(varName_i,varNames)==1){
    legend("topright",legend=MPs,lty=1:length(MPs),bty="n")
  }
  if(is.null(vline)){vline <- LHyear}
  abline(v=vline,lty=vline_lty)
}
if(is.null(title_main)){
  title_main <- paste(MSEName,MPs,sep=": ")
}
title(main=title_main,outer = TRUE)
}

```

```{r function PMsummary}
# compute and plot summary statistics for Performance metrics
PMsummary <- function(PMName,OMNames=OMNames_user,scenNames,MPs=NULL,
                      slot_to_plot="Prob", # Can be either "Prob" or "Stat"
                      main_line=-1.5,
                      hline=1,hline_lty=1,
                      par_args=list(mfrow=c(length(OMNames),1), 
                                    mar=c(0,2,0,1), oma=c(6,2,1,0),
                                    mgp=c(0.8,0.2,0), tck=-0.01),
                      ...){
PM <- get(PMName)
dots <- list(...)
if(!"ylab"%in%names(dots)){
  dots$ylab <- PMName
}

out_list <- list()

do.call(par,par_args)

for (OMNames_i in OMNames){
Names_i <- gsub("^OM_","",OMNames_i)
MSEoutNames_i = paste0(rep(gsub("OM","MSE",OMNames_i),each=length(scenNames)),
                     "_",
                     rep(scenNames,length(OMNames_i))
                     )
    out_list_i <- list()
    for(MSEoutNames_ij in MSEoutNames_i){
      scenNames_j <- str_extract(MSEoutNames_ij,"[a-z]+$")
      MSEout_ij <- get(MSEoutNames_ij)
      PMout_ij <- PM(MSEout_ij)
        MPs_ij <- PMout_ij@MPs
        out_ij <- slot(PMout_ij,slot_to_plot)
      if(!is.null(MPs)){
        out_ij <- out_ij[,match(MPs,MPs_ij)]
        MPs_ij <- MPs
      }
      dimnames(out_ij) <- list("sim"=seq(1:dim(out_ij)[1]), "MPs"=MPs_ij)
      out_ij <- cbind("scen"=scenNames_j,as.data.frame(out_ij))
      out_list_i[[MSEoutNames_ij]] <- out_ij

    }
    # Combine scenario outputs into large data frame with scen column indicating scenario
      out_df_i <- do.call(rbind,out_list_i)
      out_list[[paste0("out_df_",Names_i,"_",PMName)]] <- out_df_i
    # Plot stuff
      dots$xaxt <- ifelse(match(OMNames_i,rev(OMNames))==1,"s","n")

      do.call(boxplot,c(list(out_df_i[,-1],main=""),dots))
      mtext(Names_i,side=3,line=main_line,outer=FALSE,font=2)
      abline(h=hline,lty=hline_lty)
    }

invisible(out_list)
}
```

```{r function MSEsummary}
# compute and plot summary statistics for Performance metrics
MSEsummary <- function(MSEslotName,OMNames=OMNames_user,scenNames,MPs=NULL,
                       main_line=-1.5,
                       hline=1,hline_lty=1,
                      par_args=list(mfrow=c(length(OMNames),1),
                                    mar=c(3,2,1,1), oma=c(1,2,0,0),
                                    mgp=c(0.8,0.2,0), tck=-0.01),
                      ...){
dots <- list(...)
if(!"ylab"%in%names(dots)){
  dots$ylab <- MSEslotName
}
out_list <- list()

do.call(par,par_args)

for (OMNames_i in OMNames){
  Names_i <- gsub("^OM_","",OMNames_i)
  MSEoutNames_i = paste0(rep(gsub("OM","MSE",OMNames_i),each=length(scenNames)),
                         "_",
                         rep(scenNames,length(OMNames_i))
  )
  slot_list_i <- list()
  for(MSEoutNames_ij in MSEoutNames_i){
    scenNames_j <- str_extract(MSEoutNames_ij,"[a-z]+$")
    MSEout_ij <- get(MSEoutNames_ij)
    if(!is.expression(MSEslotName)){
      slot_k <- slot(MSEout_ij,MSEslotName) %>% apply(2,as.numeric)
    }else{
      slot_k <- eval(MSEslotName) %>% apply(2,as.numeric)
    }
    slot_k <- slot(MSEout_ij,MSEslotName) %>% apply(2,as.numeric)
    MPs_ij <- MSEout_ij@MPs
    if(!is.null(MPs)){
      slot_k <- slot_k[,match(MPs,MPs_ij)]
      MPs_ij <- MPs
    }
    dimnames(slot_k) <- list(NULL, "MPs"=MPs_ij)
    slot_k <- cbind("scen"=scenNames_j,as.data.frame(slot_k))
    slot_list_i[[MSEoutNames_ij]] <- slot_k

  }
  # Combine scenario outputs into large data frame with scen column indicating scenario
  slot_df_i <- do.call(rbind,slot_list_i)
  out_list[[paste0("slot_df_",Names_i,"_",MSEslotName)]] <- slot_df_i

  # Plot stuff
  do.call(boxplot,c(list(slot_df_i[,-1],main=""),dots))
  mtext(Names_i,side=3,line=main_line,outer=FALSE,font=2)
  abline(h=hline,lty=hline_lty)
}

invisible(out_list)
}
```

```{r data}
# Read in data for original operating models
for(file.i in list.files("OM")){
  assign(gsub(".rds","",file.i),readRDS(file.path("OM",file.i)))
}

# Read in data for modified operating models
for(file.i in list.files("OM_modified")){
  assign(gsub(".rds","",file.i),readRDS(file.path("OM_modified",file.i)))
}

if(is.null(OMNames_user)){
OMNames <- gsub(".rds","",list.files("OM"))
}else{
OMNames <- OMNames_user
}

OM_nsim <- c()
for(OM_Name_i in OMNames){
  i <- which(OMNames==OM_Name_i)
  OM_i <- get(OM_Name_i)
  nsim_i <- OM_i@nsim
  OM_nsim <- c(OM_nsim,nsim_i)
}

nsim_text <- combine_words(unique(OM_nsim),and=" or ")

# Read in MSE results
MSE_res_names <- gsub(".rds","",list.files("MSE_obj"))
#MSE_res_names <- gsub(".rds","",list.files("MSE_obj")[grepl("OM_",list.files("MSE_obj"))])
for(file.i in list.files("MSE_obj")){
  out_i <- readRDS(file.path("MSE_obj",file.i))
  names(out_i@PPD) <- out_i@MPs
  assign(gsub(".rds","",file.i),readRDS(file.path("MSE_obj",file.i)))
}
```

```{r OMNames set 1,eval=TRUE}
OMNames <- OMNames_user#[1:3]
speciesNames <- gsub("OM_","",OMNames)
speciesNames_huynh <- gsub("OM_","",OMNames_huynh)
MSEoutNames_user1 <- paste0(rep(gsub("OM","MSE",OMNames),each=length(scenNames_user1)),
                     "_",
                     rep(scenNames_user1,length(OMNames))
                     )

```

```{r plot data1, fig.cap=paste0("\\label{fig:lifeHist} Life history (growth and maturity) and vulnerability schedules at age used in the operating models during the projection period ",combine_words(speciesNames),". Growth is expressed as mean length-at-age relative to that at the maximum age."),eval=TRUE}
# Figure 1 from Huynh et al (2020)

par(mfrow=c(length(OMNames),1),mgp=c(1,0.2,0),mar=c(1.5,2,1.5,1),oma=c(1,2,1,1),tck=-0.01)
for(OM_Name_i in OMNames){
i <- which(OMNames==OM_Name_i)  
OM_i <- get(OM_Name_i)
speciesName_i <- gsub("OM_","",OM_Name_i)
plot_LH(OM_i,xlab="",ylab="",main=speciesName_i)
if(i==1){
  legend("bottomright",legend=c("Rel. length","Maturity","Vulnerability"),lwd=c(3,3,3),lty=c(1,2,3),bty="n")
}
if(i==length(OMNames)){
  mtext("Value",2,0,outer = TRUE)
}
}
```

```{r plot selectivities,fig.cap=paste0("\\label{fig:selex} Vulnerability schedules at age used in the operating models during the historical period (colored lines in rainbow order) and projection period (black lines)",combine_words(speciesNames),". Growth is expressed as mean length-at-age relative to that at the maximum age.")} 
par(mfrow=c(length(OMNames),1),mgp=c(1,0.2,0),mar=c(3,2,1.5,1),oma=c(1,2,1,1),tck=-0.01)
for(OMName_i in OMNames){#paste0(OMNames,"_base")){
OM_i <- get(OMName_i)
V_i <- OM_i@cpars$V
nyears_i <- OM_i@nyears
proyears_i <- OM_i@proyears
cols_i <- rainbow(nyears_i)
matplot(V_i[1,,1:nyears_i],type="l",lty=1,col=cols_i,main=OMName_i,xlab="age",ylab="V")
matpoints(V_i[1,,nyears_i+(1:proyears_i)],type="l",lty=1,col="black",lwd=2)
}

```

```{r plot selectivities huynh,fig.cap=paste0("\\label{fig:selexhuynh} Vulnerability schedules at age used in the operating models during the historical period (colored lines in rainbow order) and projection period (black lines)",combine_words(speciesNames_huynh),". Growth is expressed as mean length-at-age relative to that at the maximum age."),eval=FALSE} 
par(mfrow=c(length(OMNames_huynh),1),mgp=c(1,0.2,0),mar=c(3,2,1.5,1),oma=c(1,2,1,1),tck=-0.01)
for(OMName_i in OMNames_huynh){
OM_i <- get(OMName_i)
V_i <- OM_i@cpars$V
nyears_i <- OM_i@nyears
proyears_i <- OM_i@proyears
cols_i <- rainbow(nyears_i)
matplot(V_i[1,,1:nyears_i],type="l",lty=1,col=cols_i,main=OMName_i,xlab="age",ylab="V")
matpoints(V_i[1,,nyears_i+(1:proyears_i)],type="l",lty=1,col="black",lwd=2)
}

```

```{r plot Ind, eval=FALSE}
# This code is set up to make each set of plots for each MP, but the indices are basically the same

for(MP_name_i in "SCA_5"){#MPNamesAll[1]){
    # Ind
    plot_MSE_ts(
      OMNames= OMNames, scenNames=scenNames_user1,
      myvarName = expression(slot(out_i@PPD[[MP_name_i]],"Ind")),
      plot_type = "boot",
      FUN = mean_noNA,
      ylabel = "index of abundance",
      MPNamesSub = MPNamesSub_user,
      MPNamesSub_legtext=MPNamesSub_user_legtext,
      rowLabs=scenNames_key[scenNames_user1],
      oma=c(3,4,5,3),
      ylim=c(0,4)
    )
  title(main=MP_name_i,outer=TRUE)
}
```

```{r plot AddInd, eval=TRUE}
# This code is set up to make each set of plots for each MP, but the indices are basically the same
if(plot_pdf){pdf("Figs/AddInd.pdf")}
for(MP_name_i in "SCA_5"){#MPNamesAll){
    # Ind
    # MSE_i <- get(MSE_name_j)
    plot_MSE_ts(
      OMNames= OMNames, scenNames="base",
      myvarName = expression(slot(out_i@PPD[[MP_name_i]],"AddInd")),
      plot_type = "addind",
      ylabel = "index of abundance",
      # FUN=mean_noNA,
      MPNamesSub = MP_name_i,
      # MPNamesSub_legtext=MPNamesSub_user_legtext,
      #rowLabs=scenNames_key["addind"],
      xlabel="year",
      oma=c(3,4,5,3),
      ylim=c(0,8),
      main_line = -1.5,
      hline=NA
    )
  #title(main=MP_name_i,outer=TRUE)
}
if(plot_pdf){dev.off()}
```

```{r plot sims proj base, eval=FALSE}
MPs_sim <- c("SCA_1","SCA_5","iMP_avg_10")
scen <- "base"

for(OMName_i in OMNames_user){
Name_i <- gsub("^OM_","",OMName_i)
MSEoutName_i <- paste0(gsub("^OM","MSE",OMName_i),paste0("_",scen))
MSEOut_i <- get(MSEoutName_i)

figName_ij <- paste0("sims_",MSEoutName_i,"_",scen,".pdf")
if(plot_pdf){pdf(file.path("Figs",figName_ij))}
plot_MSE_sims(MSEOut_i,MPs=MPs_sim,type="proj",
              lwd=2,sims=sample(1,1:50),
              title_main = paste0(Name_i,", ",scen),
              varNames = c("Catch","TAC","FM","F_FMSY",#"recruit",
                           "SSB","SB_SBMSY"))
if(plot_pdf){dev.off()}
}

```

```{r plot sims base}
MPs_sim <- c("SCA_1","SCA_5","iMP_avg_10")
scen <- "base"

for(OMName_i in OMNames_user){
Name_i <- gsub("^OM_","",OMName_i)
MSEoutName_i <- paste0(gsub("^OM","MSE",OMName_i),paste0("_",scen))
MSEOut_i <- get(MSEoutName_i)

figName_ij <- paste0("sims_",MSEoutName_i,"_",scen,".pdf")
if(plot_pdf){pdf(file.path("Figs",figName_ij))}
plot_MSE_sims(MSEOut_i,MPs=MPs_sim,type="histproj",
              lwd=2,sims=sample(1,1:50),
              title_main = paste0(Name_i,", ",scen))
if(plot_pdf){dev.off()}
}

```

```{r plot sims minerr, eval=TRUE}
MPs_sim <- c("SCA_1","SCA_5","iMP_avg_10")
scen <- "minerr"

for(OMName_i in OMNames_user){
Name_i <- gsub("^OM_","",OMName_i)
MSEoutName_i <- paste0(gsub("^OM","MSE",OMName_i),paste0("_",scen))
MSEOut_i <- get(MSEoutName_i)

figName_ij <- paste0("sims_",MSEoutName_i,"_",scen,".pdf")
if(plot_pdf){pdf(file.path("Figs",figName_ij))}
plot_MSE_sims(MSEOut_i,MPs=MPs_sim,type="histproj",
              lwd=2,sims=1,
              title_main = paste0(Name_i,", ",scen))
if(plot_pdf){dev.off()}
}
```

```{r plot sims minrecdev}
MPs_sim <- c("SCA_1","SCA_5","iMP_avg_10")
scen <- "minrecdev"

for(OMName_i in OMNames_user){
Name_i <- gsub("^OM_","",OMName_i)
MSEoutName_i <- paste0(gsub("^OM","MSE",OMName_i),paste0("_",scen))
MSEOut_i <- get(MSEoutName_i)

figName_ij <- paste0("sims_",MSEoutName_i,"_",scen,".pdf")
if(plot_pdf){pdf(file.path("Figs",figName_ij))}
plot_MSE_sims(MSEOut_i,MPs=MPs_sim,type="histproj",
              lwd=2,sims=1,
              title_main = paste0(Name_i,", ",scen))
if(plot_pdf){dev.off()}
}

# Check if assessments are converging and hessians inverting
# unlist(lapply(MSE_BlackSeaBass_minrecdev@PPD$SCA_1@Misc,function(i){table(unlist(lapply(i$diagnostic,function(j){j$hess})))}))
```

```{r plot sims perfobs}
MPs_sim <- c("SCA_1","SCA_5","iMP_avg_10")
scen <- "perfobs"

for(OMName_i in OMNames_user){
Name_i <- gsub("^OM_","",OMName_i)
MSEoutName_i <- paste0(gsub("^OM","MSE",OMName_i),paste0("_",scen))
MSEOut_i <- get(MSEoutName_i)

figName_ij <- paste0("sims_",MSEoutName_i,"_",scen,".pdf")
if(plot_pdf){pdf(file.path("Figs",figName_ij))}
plot_MSE_sims(MSEOut_i,MPs=MPs_sim,type="histproj",
              lwd=2,sims=1,
              title_main = paste0(Name_i,", ",scen))
if(plot_pdf){dev.off()}
}

```

```{r plot sims "vtiv"}
MPs_sim <- c("SCA_1","SCA_5","iMP_avg_10")
scen <- "vtiv"

for(OMName_i in OMNames_user){
Name_i <- gsub("^OM_","",OMName_i)
MSEoutName_i <- paste0(gsub("^OM","MSE",OMName_i),paste0("_",scen))
MSEOut_i <- get(MSEoutName_i)

figName_ij <- paste0("sims_",MSEoutName_i,"_",scen,".pdf")
#if(plot_pdf){pdf(file.path("Figs",figName_ij))}
plot_MSE_sims(MSEOut_i,MPs=MPs_sim,type="histproj",
              lwd=2,sims=1,
              title_main = paste0(Name_i,", ",scen))
#if(plot_pdf){dev.off()}
}

```

```{r summary plots, eval=TRUE}
# Summary plots combining all years and all scenarios

if(plot_pdf){pdf("Figs/boxplot_P_Bgrt75BMSY.pdf")}
P75_out <- PMsummary("P75",scenNames=scenNames_user_alt,MPs=MPNamesAll,
                     cex.axis=1,las=2,outline=FALSE,
                     boxwex=0.25,ylab="",ylim=c(0,1.1),
                     hline=NA)
mtext("P(B > 0.75 BMSY)",2,outer=TRUE,font=2)
if(plot_pdf){dev.off()}

if(plot_pdf){pdf("Figs/boxplot_P_BgrtBMSY.pdf")}
P100_out <- PMsummary("P100",scenNames=scenNames_user_alt,MPs=MPNamesAll,
                     cex.axis=1,las=2,outline=FALSE,
                     boxwex=0.25,ylab="",ylim=c(0,1.1),
                     hline=NA)
mtext("P(B > BMSY)",2,outer=TRUE,font=2)
if(plot_pdf){dev.off()}

if(plot_pdf){pdf("Figs/boxplot_P_FlssFMSY.pdf")}
PNOF_out <- PMsummary("PNOF",scenNames=scenNames_user_alt,MPs=MPNamesAll,
                     cex.axis=1,las=2,outline=FALSE,
                     boxwex=0.25,ylab="",ylim=c(0,1.1),
                     hline=NA)
mtext("P(F < FMSY)",2,outer=TRUE,font=2)
if(plot_pdf){dev.off()}

if(plot_pdf){pdf("Figs/boxplot_AAVY.pdf")}
AAVY_out <- PMsummary("AAVY",scenNames=scenNames_user_alt,MPs=MPNamesAll,
                      slot_to_plot = "Stat",
                     cex.axis=1,las=2,outline=FALSE,
                     boxwex=0.25,ylab="",ylim=c(0,1.1),
                     hline=NA)
mtext("Avg. Annual Variability in Yield",2,outer=TRUE,font=2)
if(plot_pdf){dev.off()}

if(plot_pdf){pdf("Figs/boxplot_SB_SBMSY.pdf")}
MSEsummary("SB_SBMSY",scenNames=scenNames_user_alt,MPs=MPNamesAll,
           cex.axis=0.75,las=2,outline=FALSE,boxwex=0.25,ylab="")
mtext("SSB/SSBmsy",2,outer=TRUE,font=2)
if(plot_pdf){dev.off()}

if(plot_pdf){pdf("Figs/boxplot_F_FSY.pdf")}
MSEsummary("F_FMSY",scenNames=scenNames_user_alt,MPs=MPNamesAll,
           cex.axis=0.75,las=2,outline=FALSE,boxwex=0.25,ylab="")
mtext("F/Fmsy",2,outer=TRUE,font=2)
if(plot_pdf){dev.off()}

## Which is the best MP?

### Highest yield

### Most consistent yield

### Lowest probability of overfishing
```

```{r plot time series1,eval=FALSE}
# # VB
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user1,
# myvarName = "VB",
# ylabel = expression(mean~(VB)),
# FUN = mean,
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user1]
# )

# SSB
if(plot_pdf){pdf("Figs/ts_SSB.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  myvarName = "SSB",
  ylabel = expression(mean~(SSB)),
  FUN = mean,
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user1]
)
if(plot_pdf){dev.off()}
# 
# # F
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user1,  
# myvarName = "FM",
# ylabel = expression(mean~(F)),
# FUN = mean,
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user1]
# )
# 
# # TAC
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user1,  
# myvarName = "TAC",
# ylabel = expression(mean~(TAC)),
# FUN = mean,
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user1] 
# )

# Catch
if(plot_pdf){pdf("Figs/ts_Catch.pdf")}
plot_MSE_ts(
OMNames= OMNames, scenNames=scenNames_user1,  
myvarName = "Catch",
ylabel = expression(mean~(Catch)),
FUN = mean,
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user1]
)
if(plot_pdf){dev.off()}

# SSB/SSBmsy
if(plot_pdf){pdf("Figs/ts_SSB_SSBmsy.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  myvarName = "SB_SBMSY",
  ylabel = expression(mean~(SSB/SSBmsy)),
  FUN = mean,
  #ylim=c(0,2),
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user1]
)
if(plot_pdf){dev.off()}

# F/Fmsy
if(plot_pdf){pdf("Figs/ts_F_Fmsy.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  myvarName = "F_FMSY",
  ylabel = expression(mean~(F/Fmsy)),
  FUN = mean,
  #ylim=c(0,2),
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user1]
)
if(plot_pdf){dev.off()}

# C/MSY
if(plot_pdf){pdf("Figs/ts_Catch_MSY.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  myvarName = "Catch",
  myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
  ylabel = expression(mean~(Catch/MSY)),
  FUN = mean,
  #ylim=c(0,2),
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user1]
)
if(plot_pdf){dev.off()}

# 
# # TAC/MSY
# plot_MSE_ts(
#   OMNames= OMNames, scenNames=scenNames_user1,
#   myvarName = "TAC",
#   myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
#   ylabel = expression(mean~(TAC/MSY)),
#   FUN = mean,
#   ylim=c(0,2),
#   MPNamesSub = MPNamesSub_user,
#   MPNamesSub_legtext=MPNamesSub_user_legtext,
#   rowLabs=scenNames_key[scenNames_user1]
# )
# 
# # C/TAC
# plot_MSE_ts(
#   OMNames= OMNames, scenNames=scenNames_user1,  
#   myvarName = "Catch",
#   myvarDenomExpr = expression(slot(out_i,"TAC")),
#   ylabel = expression(mean~(C/TAC)),
#   FUN = mean,
#   ylim=c(0,1.1),
#   MPNamesSub = MPNamesSub_user,
#   MPNamesSub_legtext=MPNamesSub_user_legtext,
#   rowLabs=scenNames_key[scenNames_user1]
# )
# 
# # MSY
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user1,
# myvarName = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
# FUN = mean,
# ylabel= expression(mean~(MSY)),
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user1]
# )
# 
# # SSBMSY
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user1,
# myvarName = expression(slot(out_i,"RefPoint")$SSBMSY[,,-(1:out_i@nyears)]),
# FUN = mean,
# ylabel= expression(mean~(SSBMSY)),
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user1]
# )
# 
# # FMSY
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user1,
# myvarName = expression(slot(out_i,"RefPoint")$FMSY[,,-(1:out_i@nyears)]),
# FUN = mean,
# ylabel= expression(mean~(FMSY)),
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user1]
# )
# 
# # UMSY
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user1,
# myvarName = expression(F2U(slot(out_i,"RefPoint")$FMSY[,,-(1:out_i@nyears)])),
# FUN = mean,
# ylabel= expression(mean~(UMSY)),
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user1]
# )
```

```{r plot time series2,eval=TRUE}
# # VB
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user2,
# myvarName = "VB",
# ylabel = expression(mean~(VB)),
# FUN = mean,
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user2]
# )

# SSB
if(plot_pdf){pdf("Figs/ts_SSB.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user2,
  myvarName = "SSB",
  ylabel = expression(mean~(SSB)),
  FUN = mean,
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user2]
)
if(plot_pdf){dev.off()}
# 
# # F
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user2,  
# myvarName = "FM",
# ylabel = expression(mean~(F)),
# FUN = mean,
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user2]
# )
# 
# # TAC
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user2,  
# myvarName = "TAC",
# ylabel = expression(mean~(TAC)),
# FUN = mean,
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user2] 
# )

# Catch
if(plot_pdf){pdf("Figs/ts_Catch.pdf")}
plot_MSE_ts(
OMNames= OMNames, scenNames=scenNames_user2,  
myvarName = "Catch",
ylabel = expression(mean~(Catch)),
FUN = mean,
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user2]
)
if(plot_pdf){dev.off()}

# SSB/SSBmsy
if(plot_pdf){pdf("Figs/ts_SSB_SSBmsy.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user2,
  myvarName = "SB_SBMSY",
  ylabel = expression(mean~(SSB/SSBmsy)),
  FUN = mean,
  # ylim=c(0,2),
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user2]
)
if(plot_pdf){dev.off()}

# F/Fmsy
if(plot_pdf){pdf("Figs/ts_F_Fmsy.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user2,
  myvarName = "F_FMSY",
  ylabel = expression(mean~(F/Fmsy)),
  FUN = mean,
  # ylim=c(0,2),
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user2]
)
if(plot_pdf){dev.off()}

# C/MSY
if(plot_pdf){pdf("Figs/ts_Catch_MSY.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user2,
  myvarName = "Catch",
  myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
  ylabel = expression(mean~(Catch/MSY)),
  FUN = mean,
  # ylim=c(0,2),
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user2]
)
if(plot_pdf){dev.off()}

# 
# # TAC/MSY
# plot_MSE_ts(
#   OMNames= OMNames, scenNames=scenNames_user2,
#   myvarName = "TAC",
#   myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
#   ylabel = expression(mean~(TAC/MSY)),
#   FUN = mean,
#   ylim=c(0,2),
#   MPNamesSub = MPNamesSub_user,
#   MPNamesSub_legtext=MPNamesSub_user_legtext,
#   rowLabs=scenNames_key[scenNames_user2]
# )
# 
# # C/TAC
# plot_MSE_ts(
#   OMNames= OMNames, scenNames=scenNames_user2,  
#   myvarName = "Catch",
#   myvarDenomExpr = expression(slot(out_i,"TAC")),
#   ylabel = expression(mean~(C/TAC)),
#   FUN = mean,
#   ylim=c(0,1.1),
#   MPNamesSub = MPNamesSub_user,
#   MPNamesSub_legtext=MPNamesSub_user_legtext,
#   rowLabs=scenNames_key[scenNames_user2]
# )
# 
# # MSY
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user2,
# myvarName = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
# FUN = mean,
# ylabel= expression(mean~(MSY)),
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user2]
# )
# 
# # SSBMSY
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user2,
# myvarName = expression(slot(out_i,"RefPoint")$SSBMSY[,,-(1:out_i@nyears)]),
# FUN = mean,
# ylabel= expression(mean~(SSBMSY)),
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user2]
# )
# 
# # FMSY
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user2,
# myvarName = expression(slot(out_i,"RefPoint")$FMSY[,,-(1:out_i@nyears)]),
# FUN = mean,
# ylabel= expression(mean~(FMSY)),
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user2]
# )
# 
# # UMSY
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user2,
# myvarName = expression(F2U(slot(out_i,"RefPoint")$FMSY[,,-(1:out_i@nyears)])),
# FUN = mean,
# ylabel= expression(mean~(UMSY)),
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user2]
# )
```

```{r plot time series RE,eval=FALSE}
# # VB
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user1,
# myvarName = "VB",
# ylabel = expression(RE~(mean~(VB))),
# FUN = mean, MPNamesRef = "SCA_1",
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user1]
# )

# SSB
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  myvarName = "SSB",
  ylabel = expression(RE~(mean~(SBB))),
  FUN = mean, MPNamesRef = "SCA_1",
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user1],
  ylim=c(-0.5,0.5)
)
# 
# # F
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user1,  
# myvarName = "FM",
# ylabel = expression(RE~(mean~(F))),
# FUN = mean, MPNamesRef = "SCA_1",
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user1]
# )
# 
# # TAC
# plot_MSE_ts(
# OMNames= OMNames, scenNames=scenNames_user1,  
# myvarName = "TAC",
# ylabel = expression(RE~(mean~(TAC))),
# FUN = mean, MPNamesRef = "SCA_1",
# MPNamesSub = MPNamesSub_user,
# MPNamesSub_legtext=MPNamesSub_user_legtext,
# rowLabs=scenNames_key[scenNames_user1] 
# )

# Catch
plot_MSE_ts(
OMNames= OMNames, scenNames=scenNames_user1,  
myvarName = "Catch",
ylabel = expression(RE~(mean~(Catch))),
FUN = mean, MPNamesRef = "SCA_1",
MPNamesSub = MPNamesSub_user,
MPNamesSub_legtext=MPNamesSub_user_legtext,
rowLabs=scenNames_key[scenNames_user1],
ylim=c(-0.5,0.5)
)

# SSB/SSBmsy
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  myvarName = "SB_SBMSY",
  ylabel = expression(RE~(mean~(SSB/SSBmsy))),
  FUN = mean, MPNamesRef = "SCA_1",
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user1],
  ylim=c(-0.5,0.5),
  hline=NA
)

# F/Fmsy
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  myvarName = "F_FMSY",
  ylabel = expression(RE~(mean~(F/Fmsy))),
  FUN = mean, MPNamesRef = "SCA_1",
  MPNamesSub = MPNamesSub_user,
  rowLabs=scenNames_key[scenNames_user1],
  ylim=c(-1,1.5),
  hline=NA
)

# C/MSY
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  myvarName = "Catch",
  myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
  ylabel = expression(RE~(mean~(Catch/MSY))),
  FUN = mean, MPNamesRef = "SCA_1",
  MPNamesSub = MPNamesSub_user,
  MPNamesSub_legtext=MPNamesSub_user_legtext,
  rowLabs=scenNames_key[scenNames_user1],
  ylim=c(-0.5,0.5)
)

# 
# # TAC/MSY
# plot_MSE_ts(
#   OMNames= OMNames, scenNames=scenNames_user1,
#   myvarName = "TAC",
#   myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
#   ylabel = expression(RE~(mean~(TAC/MSY))),
#   FUN = mean, MPNamesRef = "SCA_1",
#   ylim=c(0,2),
#   MPNamesSub = MPNamesSub_user,
#   MPNamesSub_legtext=MPNamesSub_user_legtext,
#   rowLabs=scenNames_key[scenNames_user1],
#   ylim=c(-0.5,0.5),
#   hline=NA
# )
# 
# # C/TAC
# plot_MSE_ts(
#   OMNames= OMNames, scenNames=scenNames_user1,  
#   myvarName = "Catch",
#   myvarDenomExpr = expression(slot(out_i,"TAC")),
#   ylabel = expression(RE~(mean~(C/TAC))),
#   FUN = mean, MPNamesRef = "SCA_1",
#   ylim=c(0,1.1),
#   MPNamesSub = MPNamesSub_user,
#   MPNamesSub_legtext=MPNamesSub_user_legtext,
#   rowLabs=scenNames_key[scenNames_user1],
#   ylim=c(-0.5,0.5),
#   hline=NA
# )
# 
# # MSY
# plot_MSE_ts(
#   OMNames= OMNames, scenNames=scenNames_user1,
#   myvarName = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
#   FUN = mean, MPNamesRef = "SCA_1",
#   ylabel= expression(RE~(mean~(MSY))),
#   MPNamesSub = MPNamesSub_user,
#   MPNamesSub_legtext=MPNamesSub_user_legtext,
#   rowLabs=scenNames_key[scenNames_user1],
#   ylim=c(-0.5,0.5),
#   hline=NA
# )
# 
# # SSBMSY
# plot_MSE_ts(
#   OMNames= OMNames, scenNames=scenNames_user1,
#   myvarName = expression(slot(out_i,"RefPoint")$SSBMSY[,,-(1:out_i@nyears)]),
#   FUN = mean, MPNamesRef = "SCA_1",
#   ylabel= expression(RE~(mean~(SSBMSY))),
#   MPNamesSub = MPNamesSub_user,
#   MPNamesSub_legtext=MPNamesSub_user_legtext,
#   rowLabs=scenNames_key[scenNames_user1],
#   ylim=c(-0.5,0.5),
#   hline=NA
# )
# 
# # FMSY
# plot_MSE_ts(
#   OMNames= OMNames, scenNames=scenNames_user1,
#   myvarName = expression(slot(out_i,"RefPoint")$FMSY[,,-(1:out_i@nyears)]),
#   FUN = mean, MPNamesRef = "SCA_1",
#   ylabel= expression(RE~(mean~(FMSY))),
#   MPNamesSub = MPNamesSub_user,
#   MPNamesSub_legtext=MPNamesSub_user_legtext,
#   rowLabs=scenNames_key[scenNames_user1],
#   ylim=c(-0.5,0.5),
#   hline=NA
# )
# 
# # UMSY
# plot_MSE_ts(
#   OMNames= OMNames, scenNames=scenNames_user1,
#   myvarName = expression(F2U(slot(out_i,"RefPoint")$FMSY[,,-(1:out_i@nyears)])),
#   FUN = mean, MPNamesRef = "SCA_1",
#   ylabel= expression(RE~(mean~(UMSY))),
#   MPNamesSub = MPNamesSub_user,
#   MPNamesSub_legtext=MPNamesSub_user_legtext,
#   rowLabs=scenNames_key[scenNames_user1],
#   ylim=c(-0.5,0.5),
#   hline=NA
# )
```

```{r plot dot and whiskers, eval=FALSE}
plot_MSE_dw(
    OMNames= OMNames, scenNames=scenNames_user1,
    myvarName = "SSB",
    ylabel = expression(cv~(SSB)),
    FUN = cv,
    MPNamesSub = MPNamesAll,
    MPNamesSub_legtext=MPNamesAll_legtext,
    rowLabs=scenNames_key[scenNames_user1]
)

plot_MSE_dw(
    OMNames= OMNames, scenNames=scenNames_user1,
    myvarName = "Catch",
    myvarDenomExpr = expression(slot(out_i,"RefPoint")$MSY[,,-(1:out_i@nyears)]),
    ylabel = expression(geomean~(Catch/MSY)),
    FUN = geomean,
    MPNamesSub = MPNamesAll,
    MPNamesSub_legtext=MPNamesAll_legtext,
    rowLabs=scenNames_key[scenNames_user1],
    ylim = c(0,1)
)

```

```{r MSEtool tradeoff plots, eval=FALSE}
# MSEtool Tradeoff plots
for(outName_i in MSEoutNames_user1){
out_i <- get(outName_i)  
par(oma=c(1,1,2,1))  
MSEtool::TradePlot(out_i)
}
```

\clearpage

# Tradeoff plots

```{r tradeoffPlot1, eval=FALSE}
if(plot_pdf){pdf("Figs/tradeoffPlot1.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  # MPNamesRef = "SCA_1",
  plot_type = "tradeoff",
  PMx = Yield,
  MPNamesSub = MPNamesAll,
  MPNamesSub_legtext=MPNamesAll,
  rowLabs=scenNames_key[scenNames_user1],
  mar=c(0,0,0,0),label_axes="leftbottom",
  xlim=c(0,1.1),ylim=c(0,1.1),
  hline=NA
)
if(plot_pdf){dev.off()}
```

\clearpage

```{r tradeoffPlot2, eval=FALSE}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  # MPNamesRef = "SCA_1",
  plot_type = "tradeoff",
  PMx = PNOF,
  #proyear_phase=20,
  MPNamesSub = MPNamesAll,
  MPNamesSub_legtext=MPNamesAll,
  rowLabs=scenNames_key[scenNames_user1],
  mar=c(0,0,0,0),label_axes="leftbottom",
  xlim=c(0,1.1),ylim=c(0,1.1),
  hline=NA
)

```
\clearpage

```{r tradeoffPlot3, eval=FALSE}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  # MPNamesRef = "SCA_1",
  plot_type = "tradeoff",
  PMx = AAVY10,
  #proyear_phase=20,
  MPNamesSub = MPNamesAll,
  MPNamesSub_legtext=MPNamesAll,
  rowLabs=scenNames_key[scenNames_user1],
  mar=c(0,0,0,0),label_axes="leftbottom",
  xlim=c(0,1.1),ylim=c(0,1.1),
  hline=NA
)
```

```{r tradeoffPlot4, eval=FALSE}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  # MPNamesRef = "SCA_1",
  plot_type = "tradeoff",
  PMx = AAVY10,
  PMy = PNOF,
  #proyear_phase=20,
  MPNamesSub = MPNamesAll,
  MPNamesSub_legtext=MPNamesAll,
  rowLabs=scenNames_key[scenNames_user1],
  mar=c(0,0,0,0),label_axes="leftbottom",
  xlim=c(0,1.1),ylim=c(0,1.1),
  hline=NA
)

```

\clearpage

# Phase plots

```{r phase plots}
if(plot_pdf){pdf("Figs/phasePlot1.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user1,
  plot_type = "phase",
  #proyear_phase=50,
  MPNamesSub = c(MPNamesSub_user),
  MPNamesSub_legtext=c(MPNamesSub_user_legtext),
  rowLabs=scenNames_key[scenNames_user1],
  mar=c(0,0,0,0),label_axes="leftbottom",
  xlim=c(0,5),ylim=c(0,5),
  hline=NA
)
if(plot_pdf){dev.off()}

if(plot_pdf){pdf("Figs/phasePlot2.pdf")}
plot_MSE_ts(
  OMNames= OMNames, scenNames=scenNames_user2,
  plot_type = "phase",
  #proyear_phase=50,
  MPNamesSub = c(MPNamesSub_user),
  MPNamesSub_legtext=c(MPNamesSub_user_legtext),
  rowLabs=scenNames_key[scenNames_user2],
  mar=c(0,0,0,0),label_axes="leftbottom",
  xlim=c(0,5),ylim=c(0,5),
  hline=NA
)
if(plot_pdf){dev.off()}
```